---
title: "week_2_markdown"
author: "Tim Fässler"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

```{r}
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",") 

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

# Task 2

```{r}

wildschwein <- wildschwein_BE %>% group_by(TierName) %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs" )))

wildschwein %>% group_by(TierName) %>% summarise(mean(timelag, na.rm = TRUE))

wildschwein

summary(wildschwein$timelag)

ggplot(wildschwein) +
  geom_line(aes(DatetimeUTC, timelag, color = TierName))

ggplot(wildschwein) +
  geom_point(aes(DatetimeUTC, TierName))
  

```
Questions:

How many individuals were tracked? -> 3
For how long were the individual tracked? Are there gaps? -> Between about a half a year (Rosa) and about a year(Sabi).
Were all individuals tracked concurrently or sequentially? -> Concurrently mostly
What is the temporal sampling interval between the locations? -> The means of the timelags vary between ca. 27 min (Ruth) and ca. 21 min (Sabi). there seem to be a few outliers, especially in Ruths case, but the never are longer than ca. 17 hours so it still might not be seen as a too large of a gap.


# Task 3

```{r}
wildschwein <- wildschwein %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ))

wildschwein <- wildschwein %>% group_by(TierName) %>% mutate(speed = steplength/lead(timelag))

summary(wildschwein$speed)

wildschwein %>% group_by(TierName) %>% summarise(mean(speed, na.rm = T))
```
Questions: 
What speed unit do you get? -> The speed is in meter/second but could of course be changed to km/h. The variable (especially means of it or similar) however needs to be used with caution anyway because it sometimes is over a very large timeperiod like the 16+ hours of one timelag measure.

# Task 4

```{r}
caro <- read_delim("caro60.csv",",") 

caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)

seq3 = seq(1, 200, by = 3)
seq6 = seq(1, 200, by = 6)
seq9 = seq(1, 200, by = 9)

caro_3 = slice(caro, seq3)
caro_6 = slice(caro, seq6)
caro_9 = slice(caro, seq9)

caro <- caro %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs"))) %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ) ) %>% mutate(speed = steplength/lead(timelag))

caro_3 <- caro_3 %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs"))) %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ) ) %>% mutate(speed = steplength/lead(timelag))

caro_6 <- caro_6 %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs"))) %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ) ) %>% mutate(speed = steplength/lead(timelag))

caro_9 <- caro_9 %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs"))) %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ) ) %>% mutate(speed = steplength/lead(timelag))

colors_legend = c("1 minute intervals" = "blue", "3 minute intervals" = "green", "6 minute intervals" = "red", "9 minute intervals" = "yellow")

ggplot() +
  geom_line(caro, mapping = aes(DatetimeUTC, speed, color = "1 minute intervals")) +
  geom_line(caro_3, mapping = aes(DatetimeUTC, speed, color = "3 minute intervals")) +
  geom_line(caro_6, mapping = aes(DatetimeUTC, speed, color = "6 minute intervals")) +
  geom_line(caro_9, mapping = aes(DatetimeUTC, speed, color = "9 minute intervals")) +
  scale_color_manual(values = colors_legend)
  
ggplot() +
  geom_point(caro, mapping = aes(E, N, color = "1 minute interval")) +
  geom_path(caro, mapping = aes(E,N, color = "1 minute interval")) +
  geom_point(caro_3, mapping = aes(E, N, color = "3 minute interval")) +
  geom_path(caro_3, mapping = aes(E,N, color = "3 minute interval"))

ggplot() +
  geom_point(caro, mapping = aes(E, N, color = "1 minute interval")) +
  geom_path(caro, mapping = aes(E,N, color = "1 minute interval")) +
  geom_point(caro_6, mapping = aes(E, N, color = "6 minute interval")) +
  geom_path(caro_6, mapping = aes(E,N, color = "6 minute interval"))

ggplot() +
  geom_point(caro, mapping = aes(E, N, color = "1 minute interval")) +
  geom_path(caro, mapping = aes(E,N, color = "1 minute interval")) +
  geom_point(caro_9, mapping = aes(E, N, color = "9 minute interval")) +
  geom_path(caro_9, mapping = aes(E,N, color = "9 minute interval"))

```
Questions:

Intrepret the plots -> In the caro 3 and caro 6 the main trajectories are being preserved and even "simplified" as only the most important "staying points" of Caro remain. In the 9 minute intervals you start to lose some information like the upper part, so that interval would be to big for this data. In the first plot you see a gradual smoothing of the speed which comes naturally with the smoothing or "simplification" of the spatial pattern of the following trajectories.

