---
title: "week_2_markdown"
author: "Tim Fässler"
date: "5/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

```{r}
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",") # adjust path

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

# Task 2

```{r}

wildschwein <- wildschwein_BE %>% group_by(TierName) %>% mutate(timelag = as.integer(difftime(DatetimeUTC, lag(DatetimeUTC), units = "secs" )))

wildschwein %>% group_by(TierName) %>% summarise(mean(timelag, na.rm = TRUE))

wildschwein

summary(wildschwein$timelag)

ggplot(wildschwein) +
  geom_line(aes(DatetimeUTC, timelag, color = TierName))

ggplot(wildschwein) +
  geom_point(aes(DatetimeUTC, TierName))
  

```
Questions:

How many individuals were tracked? -> 3
For how long were the individual tracked? Are there gaps? -> Between about a half a year (Rosa) and about a year(Sabi).
Were all individuals tracked concurrently or sequentially? -> Concurrently mostly
What is the temporal sampling interval between the locations? -> The means of the timelags vary between ca. 27 min (Ruth) and ca. 21 min (Sabi). there seem to be a few outliers, especially in Ruths case, but the never are longer than ca. 17 hours so it still might not be seen as a too large of a gap.


# Task 3

```{r}
wildschwein <- wildschwein %>% mutate(steplength = sqrt( ((E - lead(E, 1))^2) + ((N - lead(N, 1))^2) ))

wildschwein <- wildschwein %>% group_by(TierName) %>% mutate(speed = steplength/lead(timelag))

summary(wildschwein$speed)

wildschwein %>% group_by(TierName) %>% summarise(mean(speed, na.rm = T))
```
Questions: 
What speed unit do you get? -> The speed is in meter/second but could of course be changed to km/h. The variable (especially means of it or similar) however needs to be used with caution anyway because it sometimes is over a very large timeperiod like the 16+ hours of one timelag measure.


